---
title: "Using rucrdtw"
author: "Philipp Boersch-Supan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using rucrdtw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: rucrdtw.bib
---

`rucrdtw` provides R bindings for functions from the [UCR Suite](http://www.cs.ucr.edu/~eamonn/UCRsuite.html) [@rakthanmanon2012searching], which enables ultrafast subsequence search under both Dynamic Time Warping and Euclidean Distance.

## Installation
Install `rucrdtw` from GitHub:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("pboesu/rucrdtw")
```

## Examples

Load `rucrdtw` package:
```{r}
library("rucrdtw")
```


create a random long time series
```{r random-walk}
set.seed(123)
rwalk <- cumsum(runif(1e7, min = -0.5, max = 0.5))
```
Pick a random subsequence of 100 elements as a query
```{r rw-query}
qstart <- sample(length(rwalk), 1)
query <- rwalk[qstart:(qstart+100)]
```
Since both query and data are R vectors, we use the vector-vector methods for the search.
```{r rw-search}
system.time(dtw_search <- ucrdtw_vv(data = rwalk, query = query, dtwwindow = 0.05))
all.equal(qstart, dtw_search$location)
system.time(ed_search <- ucred_vv(data = rwalk, query = query))
all.equal(qstart, ed_search$location)
```
And in a matter of seconds we have searched 10 million data points and rediscovered our query!

Searching for an exact match, however, is somewhat artificial. The real power of the similarity search is finding structurally similar subsequences in complex sets of time series. To demonstarte this we load an example data set:

```{r load-data}
data("synthetic_control")
```

This data set contains 600 time series of length 60 from 6 classes [@alcock1999time]. The data set documentation contains further information about these data. It can be displayed using the command `?synthetic_control`. We can plot an example of each class 

```{r plot-examples, fig.width=6}
par(mfrow = c(3,2),
    mar = c(1,1,1,1))
classes = c("Normal", "Cyclic", "Increasing", "Decreasing", "Upward shift", "Downward shift")
for (i in 1:6){
  plot(synthetic_control[i*100-99,], type = "l", xaxt = "n", yaxt = "n", ylab="", xlab = "", bty="n", main=classes[i])
}
```

Since we are now comparing a query against a set of time series, we only need to do comparisons for non-overlapping data sequences. The matrix-vector methods `ucrdtw_mv` and `ucred_mv` provide this functionality.

We can demonstrate this by removing a query from the data set, and then searching for a closest match:
```{r, echo=TRUE, message=TRUE, warning=TRUE}
index <- 600
query <- synthetic_control[index,]
#microbenchmark::microbenchmark(
dtw_search = ucrdtw_mv(synthetic_control[-index,], query, 0.05, byrow = TRUE)
ed_search = ucred_mv(synthetic_control[-index,], query, byrow= TRUE)
#times=50)
```
And plot the results:
```{r plot-search, fig.width=6}
plot(synthetic_control[dtw_search$location,], type="l", ylim=c(0,55), ylab="")
lines(query, col="red")
lines(synthetic_control[ed_search$location,], col="blue", lty=3, lwd=3)
legend("topright", legend = c("query", "DTW match", "ED match"), col=c("red", "black", "blue"), lty=c(1,1,3), bty="n")
```

## References
