---
title: "Using rucrdtw"
author: "Philipp Boersch-Supan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using rucrdtw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: rucrdtw.bib
---

`rucrdtw` provides R bindings for functions from the [UCR Suite](http://www.cs.ucr.edu/~eamonn/UCRsuite.html) [@rakthanmanon2012searching], which enables ultrafast subsequence search under both Dynamic Time Warping and Euclidean Distance.

## Installation
Install `rucrdtw` from GitHub:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("pboesu/rucrdtw")
```

## Examples

Load `rucrdtw` package:
```{r}
library("rucrdtw")
```

Load the example data set.

```{r load-data}
data("synthetic_control")
```

This data set contains 600 time series of length 60 from 6 classes [@alcock1999time]. The data set documentation contains further information about these data. It can be displayed using the command `?synthetic_control`. We can plot an example of each class 

```{r plot-examples, fig.width=6}
par(mfrow = c(3,2),
    mar = c(1,1,1,1))
classes = c("Normal", "Cyclic", "Increasing", "Decreasing", "Upward shift", "Downward shift")
for (i in 1:6){
  plot(synthetic_control[i*100-99,], type = "l", xaxt = "n", yaxt = "n", ylab="", xlab = "", bty="n", main=classes[i])
}
```

`rucrdtw` provides tools to efficiently detect a pattern such as the examples given above in a longer time series. We can demonstrate this, by pasting together the 600 example series, and searching for a predetermined subsequence.

```{r search-exact, fig.width=6}
#reshape matrix into vector
long_series <- as.vector(t(synthetic_control))
#pick series 175 as query
query <- synthetic_control[175,]
#calculate starting location in the long series
query_start <- 175*ncol(synthetic_control)-59

#plot long series
plot(long_series, type = "l")
#overlay known position of query
lines(seq(query_start, by=1, length.out = 60), synthetic_control[175,], col="red")

#do the search using the vector-vector method
dtw_search <- ucrdtw_vv(long_series, query, length(query), 0.05)

#visualize the result
plot(long_series, type = "l", xlim = c(query_start-100, query_start+160))
abline(v=c(dtw_search$location, dtw_search$location+length(query)) , col="blue")
lines(seq(query_start, by=1, length.out = 60), synthetic_control[175,], col="red")
```
