---
title: "Using rucrdtw"
author: "Philipp Boersch-Supan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using rucrdtw}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: rucrdtw.bib
---

`rucrdtw` provides R bindings for functions from the [UCR Suite](http://www.cs.ucr.edu/~eamonn/UCRsuite.html) [@rakthanmanon2012searching], which enables ultrafast subsequence search under both Dynamic Time Warping and Euclidean Distance.

## Installation
Install `rucrdtw` from GitHub:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("pboesu/rucrdtw")
```

## Examples

Load `rucrdtw` package:
```{r}
library("rucrdtw")
```

Load the example data set.

```{r load-data}
data("synthetic_control")
```

This data set contains 600 time series of length 60 from 6 classes [@alcock1999time]. The data set documentation contains further information about these data. It can be displayed using the command `?synthetic_control`. We can plot an example of each class 

```{r plot-examples, fig.width=6}
par(mfrow = c(3,2),
    mar = c(1,1,1,1))
classes = c("Normal", "Cyclic", "Increasing", "Decreasing", "Upward shift", "Downward shift")
for (i in 1:6){
  plot(synthetic_control[i*100-99,], type = "l", xaxt = "n", yaxt = "n", ylab="", xlab = "", bty="n", main=classes[i])
}
```

`rucrdtw` provides tools to efficiently detect a pattern such as the examples given above in a longer time series. We can demonstrate this, by pasting together the 600 example series, and searching for a predetermined subsequence.

```{r search-exact1, fig.width=6}
#reshape matrix into vector
long_series <- as.vector(t(synthetic_control))
#pick series 175 as query
query <- synthetic_control[175,]
#calculate starting location in the long series
query_start <- 175*ncol(synthetic_control)-59

#plot long series
plot(long_series, type = "l")
#overlay known position of query
lines(seq(query_start, by=1, length.out = 60), synthetic_control[175,], col="red")
```

We can then search using both DTW and Euclidean distance measures.

```{r search-exact2}
#do the search using the vector-vector method
dtw_search <- ucrdtw_vv(long_series, query, length(query), 0.05)
ed_search <- ucred_vv(long_series, query, length(query))
```

And plot the search result onto the data.
```{r search-results, fig.width=6}
#visualize the result
plot(long_series, type = "l", xlim = c(query_start-100, query_start+160))
abline(v=c(dtw_search$location, dtw_search$location+length(query)) , col="blue")
abline(v=c(ed_search$location, ed_search$location+length(query)) , col="darkgreen", lty=3, lwd=3)
lines(seq(query_start, by=1, length.out = 60), synthetic_control[175,], col="red")
```

Searching for an exact match, however, doesn't require an expensive similarity measure such as DTW. The real power, however, of the DTW similarity search, is finding structurally similar subsequences in complex sets of time series.

We can demonstrate this by removing the query from the data set:
```{r, echo=TRUE, message=TRUE, warning=TRUE}
index <- 600
query <- synthetic_control[index,]
#microbenchmark::microbenchmark(
dtw_search = ucrdtw_vm(synthetic_control[-index,], query, length(query), 0.05, byrow = TRUE)
ed_search = ucred_vm(synthetic_control[-index,], query, length(query), byrow= TRUE)
#times=50)
plot(synthetic_control[dtw_search$location,], type="l", ylim=c(0,55))
lines(query, col="red", lty=2)
lines(synthetic_control[ed_search$location,], col="blue", lty=3, lwd=3)
```

